---
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
params:
  data_frame: NA
  resistant_level: "resistant"
  control_level: "control"
  treatment1_level: NA
  n_dec: 3
  model_type: "lmm"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA)

library(dplyr)
library(stringr)
library(lme4)
library(purrr)
library(tidyr)
library(ggplot2)
library(tidyselect)
library(DT)
library(kableExtra)
library(multcomp)
library(forcats)
```


```{r data}
data_frame <- params$data_frame |> 
  mutate(across(1:3, factor))
df_names <- names(data_frame)

data_model <- data_frame |>
  mutate(
    "{df_names[2]}" := .data[[df_names[2]]] |>
      fct_relevel(params$resistant_level),
    "{df_names[3]}" := .data[[df_names[3]]] |>
      fct_relevel(params$control_level)
  )

# If treatment has 3 levels, treatrment1 is located after control
if (!is.na(params$treatment1_level)) {
  data_model <- data_model |> 
    mutate(
      "{df_names[3]}" := .data[[df_names[3]]] |>
      fct_relevel(params$treatment1_level, after = 1L)
    )
}


if (params$model_type == "glmm") {
  data_model <- data_frame |>
    mutate(
      "{df_names[4]}" := .data[[df_names[4]]] / 100,
      weight_100 = 100
    )
}

```

---
title: "`r str_c(str_to_upper(params$model_type), ": ", df_names[4], " ~ ", df_names[2], " x ", df_names[3])`"
---

```{r model}
# Formulas
# With random intercepts only
form1 <- call(
  "formula",
  str_c(
    df_names[4], " ~ ", df_names[2], " * ", df_names[3],
    " + (1 | ", df_names[1], ")"
  )
)
# With random intercepts and slopes
form2 <- call(
  "formula",
  str_c(
    df_names[4], " ~ ", df_names[2], " * ", df_names[3],
    " + (1 + ", df_names[3], " | ", df_names[1], ")"
  )
)


if (params$model_type == "lmm") {
  model1 <- lmer(eval(form1), data = data_model)
  model2 <- lmer(eval(form2), data = data_model)
}

if (params$model_type == "glmm") {
  model1 <- glmer(
    eval(form1), data = data_model, family = "binomial", weights = weight_100
  )
  model2 <- glmer(
    eval(form2), data = data_model, family = "binomial", weights = weight_100
  )
}

# Model comparison, when running a LMM, refiting from REML to ML is not needed
# since the fixed part is the same in both models.
# (https://stats.stackexchange.com/questions/465879/different-results-for-model-comparison-using-anova-function-for-reml-and-ml)
comp_models <- anova(model1, model2, refit = FALSE)

# If both models are different, we choose the best. If not, we choose the less
# complex (random intercepts only)
if (comp_models$`Pr(>Chisq)`[2] <= 0.05) {
  # The best model is allways on the second row.
  model <- get(rownames(comp_models)[2])
} else {
  model <- model1
}

```

```{r multcomp}

# The data_frame passed to this file MUST have...:
# - The sensitivity factor in the second column and the resistant level must be
#   the first one.
# - The treatment factor in the third column and the control level must be the 
#   first one.
sensitive_level <- levels(data_model[[2]])[2]


if (nlevels(data_model[[3]]) == 2) {
  treatment_level <- levels(data_model[[3]])[2]
  
  ContSen_ContRes <- str_c(
    "[", params$control_level, " & ", sensitive_level, "] - [",
    params$control_level, " & ", params$resistant_level, "]" 
  )
  TrtRes_ContRes <- str_c(
    "[", treatment_level, " & ", params$resistant_level, "] - [",
    params$control_level, " & ", params$resistant_level, "]"
  )
  TrtSen_ContSen <- str_c(
    "[", treatment_level, " & ", sensitive_level, "] - [", 
    params$control_level, " & ", sensitive_level, "]" 
  )
  interaction <- str_c(df_names[2], ":", df_names[3], " interaction")
  
  
  K <- rbind(
    "ContSen - ContRes" = c(0, 1, 0, 0),
    "TrtRes - ContRes"  = c(0, 0, 1, 0),
    "TrtSen - ContSen"  = c(0, 0, 1, 1),
    "Interaction"       = c(0, 0, 0, 1)
  )
  
  row.names(K) <- c(
    ContSen_ContRes, TrtRes_ContRes, TrtSen_ContSen, interaction
  )
  
} else {
  treatment2_level <- levels(data_model[[3]])[3]
  
  ContSen_ContRes <- str_c(
    "[",params$control_level, " & ", sensitive_level, "] - [",
    params$control_level, " & ", params$resistant_level, "]"
  )
  TrtRes_ContRes <- str_c(
    "[",params$treatment1_level, " & ", params$resistant_level, "] - [",
     params$control_level, " & ", params$resistant_level, "]"
  )
  TrtSen_ContSen <- str_c(
    "[", params$treatment1_level, " & ", sensitive_level, "] - [",
     params$control_level, " & ", sensitive_level, "]"
  )
  Trt2Res_ContRes <- str_c(
    "[", treatment2_level, " & ", params$resistant_level, "] - [",
    params$control_level, " & ", params$resistant_level, "]"
  )
  Trt2Sen_ContSen <- str_c(
    "[", treatment2_level, " & ", sensitive_level, "] - [",
    params$control_level, " & ", sensitive_level, "]"
  )
  sensitivity_Trt_interaction <-  str_c(
    df_names[2], ":", params$treatment1_level, " interaction"
  )
  sensitivity_Trt2_interaction <-  str_c(
    df_names[2], ":", treatment2_level, " interaction"
  )
  difference_interactions <- str_c(
    "[", df_names[2], ":", treatment2_level, "] - [",
     df_names[2], ":", params$treatment1_level, "]"
  )
  Trt2Res_TrtRes <- str_c(
    "[", treatment2_level, " & ", params$resistant_level, "] - [",
    params$treatment1_level, " & ", params$resistant_level, "]"
  )
  Trt2Sen_TrtSen <- str_c(
    "[", treatment2_level, " & ", sensitive_level, "] - [",
    params$treatment1_level, " & ", sensitive_level, "]"
  )
  
  K <- rbind(
    "ContSen - ContRes"    = c(0, 1, 0, 0, 0, 0),
    "TrtRes - ContRes"     = c(0, 0, 1, 0, 0, 0),
    "TrtSen - ContSen"     = c(0, 0, 1, 0, 1, 0),
    "Trt2Res - ContRes"    = c(0, 0, 0, 1, 0, 0),
    "Trt2Sen - ContSen"    = c(0, 0, 0, 1, 0, 1),
    "sensitivity:Trt"      = c(0, 0, 0, 0, 1, 0),
    "sensitivity:Trt2"     = c(0, 0, 0, 0, 0, 1),
    "sens:Trt2 - sens:Trt" = c(0, 0, 0, 0,-1, 1),
    "Trt2Res - TrtRes"     = c(0, 0,-1, 1, 0, 0),
    "Trt2Sen - TrtSen"     = c(0, 0,-1, 1,-1, 1)
  )
  
  row.names(K) <- c(
    ContSen_ContRes, 
    TrtRes_ContRes, TrtSen_ContSen, 
    Trt2Res_ContRes, Trt2Sen_ContSen,
    sensitivity_Trt_interaction, sensitivity_Trt2_interaction,
    difference_interactions, Trt2Res_TrtRes, Trt2Sen_TrtSen
  )
  
}

summary <- glht(model, K) |> summary()

result <- tibble(
 Hypothesis = str_c(names(summary$test$coefficients), " = 0"),
 Estimate = summary$test$coefficients,
 `Std. Error` = summary$test$sigma,
 `z value` = summary$test$tstat,
 `p-value` = summary$test$pvalues
) |> 
  mutate(
    across(where(is.numeric), ~round(., n_dec))
  )

```


## Data

### Original Data

```{r data-table}
data_frame |> 
  mutate(
    across(1:3, factor)
  ) |> 
  datatable(data_frame, filter = "top")

```

### Data Structure

```{r data-structure}
rnd_fct <- str_c(
  str_c("__", df_names[1], "__"), "with",
  unique(data_frame[[1]]) |> length(),
  "subjects", sep = " "
)

sen_fct <- str_c(
  str_c("__", df_names[2], "__"), "with levels",
  str_c(
    str_c("__", levels(data_model[[2]]), "__"),
    c(" (resistant level)", " (sensible level)"),
    collapse = " and "
  ),
  sep = " "
)

if (nlevels(data_model[[3]]) == 2) {
  
  trt_fct <- str_c(
    str_c("__", df_names[3], "__"), "with levels",
    str_c(
      str_c("__", levels(data_model[[3]]), "__"),
      c(" (control level)", " (treatment level)"),
      collapse = " and "
    ),
    sep = " "
  )
  
} else {
  
   trt_fct <- str_c(
    str_c("__", df_names[3], "__"), "with levels",
    str_c(
      str_c("__", levels(data_model[[3]]), "__"),
      c(" (control level)", " (treatment 1 level)",  " (treatment 2 level)"),
      collapse = " and "
    ),
    sep = " "
  ) |> 
     str_replace(" and", ",")
  
}

resp <- str_c(
  str_c("__", df_names[4], "__"), sep = " "
)


```

- **PDX** random factor: `r rnd_fct`
- **Sensitivity** fixed factor: `r sen_fct`
- **Treatment** fixed factor: `r trt_fct`
- **Response** variable: `r resp`

```{r data-count}

count_data <- left_join(
  data_model |>
    dplyr::select(1:3) |> 
    unique() |> 
    count(.data[[df_names[2]]], .data[[df_names[3]]], name = "n PDX"),
  data_model |> 
    count(.data[[df_names[2]]], .data[[df_names[3]]], name = "n Measures")
) |> 
  suppressMessages()

kbl(count_data) |> 
  kable_styling(full_width = FALSE)

```


## Graphic

```{r graphic}

paired_data <- data_model |> 
  group_by(.data[[df_names[1]]], .data[[df_names[2]]], .data[[df_names[3]]]) |>
  summarise(
    "{df_names[4]}" := mean(.data[[df_names[4]]]),
    .groups = "drop"
  )

ggplot(paired_data, aes(.data[[df_names[3]]], .data[[df_names[4]]])) +
  facet_wrap(~ .data[[df_names[2]]]) +
  geom_point(cex = 2) +
  geom_line(aes(group = .data[[df_names[1]]])) +
  theme_classic() +
  theme(strip.background = element_blank(), text=element_text(size=20))

# run next graphic only if 3 levels treatment

eval_graph2 <- !is.na(treatment1_level)

```

```{r graphic2, eval=eval_graph2, fig.asp=1.25}

trt1_plot <- paired_data |>
  mutate(
    control_treatment = case_when(
      group == levels(.data[[df_names[3]]])[1] ~ "Control",
      TRUE ~ "Treatment"
    ),
    plot_treatment = levels(.data[[df_names[3]]])[2]
  ) |>
  filter(
    .data[[df_names[3]]] == levels(.data[[df_names[3]]])[1] |
    .data[[df_names[3]]] == levels(.data[[df_names[3]]])[2]
  )


trt2_plot <- paired_data |>
  mutate(
    control_treatment = case_when(
      group == levels(.data[[df_names[3]]])[1] ~ "Control",
      TRUE ~ "Treatment"
    ),
    plot_treatment = levels(.data[[df_names[3]]])[3]
  ) |>
  filter(
    .data[[df_names[3]]] == levels(.data[[df_names[3]]])[1] |
    .data[[df_names[3]]] == levels(.data[[df_names[3]]])[3]
  )


bind_rows(trt1_plot, trt2_plot) |>
  mutate(
    plot_treatment = factor(plot_treatment) |> 
      fct_relevel(treatment1_level)
  ) |> 
  ggplot(aes(control_treatment, .data[[df_names[4]]])) +
  facet_grid(plot_treatment ~ .data[[df_names[2]]]) +
  geom_point(cex = 2) +
  geom_line(aes(group = .data[[df_names[1]]])) +
  theme(
    strip.background = element_blank(), text=element_text(size=20)
  ) + 
  labs(x = "")
```



## Result

### Hypothesis Testing
```{r multcomp-tbl}

kbl(result) |> 
  kable_styling(full_width = FALSE)
 
```

### Model Summary
```{r model-summary}

summary(model)

```


## References

### R
```{r ref-r}
citation()
```

### lme4
Package used to perform the `r str_to_upper(params$model_type)` model.
```{r ref-lme4}
citation("lme4")
```

### multcomp
Package used for simultaneous hypothesis testing.
```{r ref-multcomp}
citation("multcomp")
```

## Session Info

```{r session}
sessionInfo()
```




